---
title: "Phenograph Median Correlation Mapping"
output:
  html_document:
    df_print: paged
---

In this notebook, I'll take results from 02_Jaccard_tSNE, perform phenograph clustering of samples to get neighborhoods of related cells, and group those cells to generate TSS-centric counts for mapping to transcriptomic types using a simple median correlation.

First, load necessary libraries:
```{r Load libraries}
library(dplyr)
library(matrixStats)
library(lowcat)
library(ggplot2)
library(GenomicRanges)
library(Rphenograph)
library(purrr)
options(stringsAsFactors = F)
```

Median matrices were already computed for subclasses and clusters in 01_Ingest_and_QC. We'll also need to select genes to use for this analysis based on comonality between gene expression data and TSS regions, and a sd cutoff for variance:
```{r Load medians}
#sd_cutoff <- 2
#load(file.path(pdir,"V1.cl.rda"))
load("select.markers.rda")

load("../common/tss_regions_2e4.rda")
load("cl.med.rda")
load("sub.med.rda")

subclass_med_mat <- sub.med[select.markers,]
cluster_med_mat <- cl.med[select.markers,]

# subclass_med_mat <- matrix_var_filter(subclass_med_mat, sd_cut = sd_cutoff)
# cluster_med_mat <- matrix_var_filter(cluster_med_mat, sd_cut = sd_cutoff)

subclass_tss_regions <- tss_regions %>%
  filter(name %in% rownames(subclass_med_mat))

subclass_tss_regions_GR <- bed_to_GRanges(subclass_tss_regions)

cluster_tss_regions <- tss_regions %>%
  filter(name %in% rownames(cluster_med_mat))

cluster_tss_regions_GR <- bed_to_GRanges(cluster_tss_regions)

```

We also need to read annotations from the same .tome that was used to build the median matrices:
```{r Load annotations}
load("anno.rda")

subclass_anno <- anno %>%
  select(subclass_id, subclass_label, subclass_color) %>%
  unique()
names(subclass_anno) <- paste0("pred_",names(subclass_anno))

cluster_anno <- anno %>%
  select(cluster_id, cluster_label, cluster_color) %>%
  unique()
names(cluster_anno) <- paste0("pred_",names(cluster_anno))

```

We'll also need the tSNE coordinates and fragments from the BAM files:
```{r Load scATAC}
# Load Jaccard and tSNE results
load("f1e4_e25_c10_ds1e4_x1e3_jaccard.rda")

# load bam fragments
load("f1e4_e25_c10_bam_fragments.rda")

bam_fragments <- bam_fragments[samples$sample_id]
bam_fragments <- bam_fragments[names(bam_fragments) %in% samples$sample_id]
```

Perform Rphenograph clustering on tSNE coordinate space:
```{r}
tsne_mat <- as.matrix(tsne_df[,c("x","y")])

pg <- Rphenograph(tsne_mat, k = 6)
pg_clust <- as.numeric(membership(pg[[2]]))

pg_table <- as.data.frame(table(membership(pg[[2]])))
pg_table

#hist(pg_table, breaks = 20)
```

Plot Rphenograph clustering: 
```{r}
tsne_df$pg_clust <- pg_clust
tsne_df$pg_clust_color <- varibow(max(pg_clust))[pg_clust]

ggplot() + geom_point(data = tsne_df,
                      aes(x = x, y = y,
                          color = pg_clust_color)) +
  scale_color_identity()

```

Let's balance the contributions of each cell within each cluster by sampling an equal number of reads based on the member with the lowest number of reads:
```{r}
pg_clusters <- unique(tsne_df$pg_clust)

balanced_fragments <- balance_fragment_clusters(bam_fragments,
                                                 pg_clusters)

```


Now, we'll compute counts for each sample vs each of these region sets:
```{r}
subclass_tss_counts <- region_fragment_count_clusters(fragment_list = balanced_fragments,
                                                      fragment_clusters = tsne_df$pg_clust,
                                                      subclass_tss_regions_GR)

cluster_tss_counts <- region_fragment_count_clusters(fragment_list = balanced_fragments,
                                                     fragment_clusters = tsne_df$pg_clust,
                                                     cluster_tss_regions_GR)
```

With counts, we can run correlations with the columns in the subclass and cluster matrices:
```{r}
# This is a key step!
common_subclass_med <- subclass_med_mat[rownames(subclass_tss_counts),]
# Above is a key step!
subclass_max_cor <- max_column_correlation(subclass_tss_counts,
                                           log2(common_subclass_med + 1))
names(subclass_max_cor) <- c("pg_clust","pred_subclass_cor","pred_subclass_id")
subclass_max_cor$pg_clust <- as.numeric(subclass_max_cor$pg_clust)
subclass_max_cor$pred_subclass_id <- as.numeric(subclass_max_cor$pred_subclass_id)

common_cluster_med <- cluster_med_mat[rownames(cluster_tss_counts),]
cluster_max_cor <- max_column_correlation(cluster_tss_counts,
                                          log2(common_cluster_med + 1))
names(cluster_max_cor) <- c("pg_clust","pred_cluster_cor","pred_cluster_id")
cluster_max_cor$pg_clust <- as.numeric(cluster_max_cor$pg_clust)
cluster_max_cor$pred_cluster_id <- as.numeric(cluster_max_cor$pred_cluster_id)
```

Let's join things together and try plotting:
```{r}
tsne_df <- tsne_df %>%
  mutate(study = ifelse(grepl("^sample_",sample_id),"Cusanovich","Graybuck")) %>%
  mutate(pch = ifelse(study == "Cusanovich", 1, 21))

subclass_tsne_data <- tsne_df %>%
  left_join(subclass_max_cor) %>%
  left_join(subclass_anno)

cluster_tsne_data <- tsne_df %>%
  left_join(cluster_max_cor) %>%
  left_join(cluster_anno)

ggplot() +
  geom_point(data = subclass_tsne_data %>%
               filter(study == "Cusanovich"),
             aes(x = x, y = y,
                 color = pred_subclass_color),
             shape = 16) +
  geom_point(data = subclass_tsne_data %>%
               filter(study == "Graybuck"),
             aes(x = x, y = y,
                 color = pred_subclass_color),
             shape = 2) +
  scale_color_identity()

ggplot() +
    geom_point(data = cluster_tsne_data %>%
               filter(study == "Cusanovich"),
             aes(x = x, y = y,
                 color = pred_cluster_color),
             shape = 16) +
  geom_point(data = cluster_tsne_data %>%
               filter(study == "Graybuck"),
             aes(x = x, y = y,
                 color = pred_cluster_color),
             shape = 2) +
  scale_color_identity()

ggplot() +
    geom_point(data = cluster_tsne_data %>%
               filter(study == "Cusanovich"),
             aes(x = x, y = y,;
                 color = as.factor(grepl("ALM",pred_cluster_label))),
             shape = 16) +
  geom_point(data = cluster_tsne_data %>%
               filter(study == "Graybuck"),
             aes(x = x, y = y,
                 color = as.factor(grepl("ALM",pred_cluster_label))),
             shape = 2)

```

```{r}
library(rbokeh)

cluster_bokeh <- figure(height = 800, width = 800) %>%
  ly_points(data = cluster_tsne_data,
            x = x, y = y,
            color = pred_cluster_color,
            hover = list("Cluster" = pred_cluster_label))
htmlwidgets::saveWidget(cluster_bokeh, "cluster_assignments.html")

subclass_bokeh <- figure(height = 800, width = 800) %>%
  ly_points(data = subclass_tsne_data,
            x = x, y = y,
            color = pred_subclass_color,
            hover = list("Subclass" = pred_subclass_label))
htmlwidgets::saveWidget(subclass_bokeh, "subclass_assignments.html")


```

Tons of cells seem to be along a single continuum for IT samples. Let's try umap to visualize:
```{r}
library(umap)

umap_res <- umap(jaccard_pcs, min_dist = 0.5)

umap_coords <- as.data.frame(umap_res$layout)
names(umap_coords) <- c("ux","uy")

cluster_umap_data <- cbind(cluster_tsne_data, umap_coords)

ggplot() +
  geom_point(data = cluster_umap_data,
             aes(x = ux, y = uy,
                 color = pred_cluster_color)) +
  scale_color_identity()

umap_bokeh <- figure(height = 800, width = 800) %>%
  ly_points(data = cluster_umap_data,
            x = ux, y = uy,
            color = pred_cluster_color,
            hover = list("Cluster" = pred_cluster_label))
htmlwidgets::saveWidget(umap_bokeh, "umap_cluster_assignments.html")

```


Save the outputs for use later:
```{r}
results <- tsne_df %>%
  left_join(subclass_max_cor) %>%
  left_join(subclass_anno) %>%
  left_join(cluster_max_cor) %>%
  left_join(cluster_anno) %>%
  cbind(umap_coords)

write.csv(results, "2e4_tss_phenograph_scatac_visp_marker_correlation.csv")

save(subclass_tss_counts,
     cluster_tss_counts,
     file = "2e4_tss_phenograph_scatac_visp_marker_correlation_matrices.rda")
```


```{r Merge BAMS}
library(purrr)

bam_dir <- "07_PMC_markers_bam"

dir.create(bam_dir)

pred_clusters <- unique(results$pred_cluster_label)

walk(pred_clusters,
     function(x) {
       cluster_samples <- results$sample_id[results$pred_cluster_label == x]
       bam_files <- samples$bam_file[match(cluster_samples, samples$sample_id)]
       x <- gsub("/","_",x)
       bam_out <- paste0(bam_dir,"/",
                         gsub(" ","_",x),
                         ".bam")
       if(!file.exists(bam_out)) {
         merge_bam_files(bam_files,
                         bam_out)
       } else {
         print(paste0(bam_out," already exists. Skipping."))
       }

     })

```

Per-subclass BAMs
```{r Merge BAMS}
library(purrr)

bam_dir <- "07_PMC_markers_subclass_bam"

dir.create(bam_dir)

pred_subclasss <- unique(results$pred_subclass_label)

walk(pred_subclasss,
     function(x) {
       subclass_samples <- results$sample_id[results$pred_subclass_label == x]
       bam_files <- samples$bam_file[match(subclass_samples, samples$sample_id)]
       x <- gsub("/","_",x)
       bam_out <- paste0(bam_dir,"/",
                         gsub(" ","_",x),
                         ".bam")
       if(!file.exists(bam_out)) {
         merge_bam_files(bam_files,
                         bam_out)
       } else {
         print(paste0(bam_out," already exists. Skipping."))
       }

     })

```

Per-plot group BAMs
```{r Merge group BAMs}
library(purrr)

bam_dir <- "07_PMC_markers_group_bam"

dir.create(bam_dir)

group_anno <- read.csv(file.path(data_dir,"cluster_grouping_for_tracks.csv"))

results <- results %>%
  left_join(group_anno)

pred_groups <- unique(results$group_label)

walk(pred_groups,
     function(x) {
       group_samples <- results$sample_id[results$group_label == x]
       bam_files <- samples$bam_file[match(group_samples, samples$sample_id)]
       x <- gsub("/","_",x)
       bam_out <- paste0(bam_dir,"/",
                         gsub(" ","_",x),
                         ".bam")
       if(!file.exists(bam_out)) {
         merge_bam_files(bam_files,
                         bam_out)
       } else {
         print(paste0(bam_out," already exists. Skipping."))
       }

     })

```